---
title: "Practical Machine Learning Course Project"
author: "Vernon Rothery"
date: "6 August 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
set.seed(1111)

```

## R Markdown

 <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars, echo=TRUE}
summary(cars)
```

## Including Plots
##Downloading from the Data Source.
Copy hyperlinks from coursera website.
Creates trainUrl and testUrl environment variables to be used to create the training and testing data files.
```{r}

trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))

testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))

inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)

myTraining <- training[inTrain, ]
myTesting <- training[-inTrain, ]
dim(myTraining); dim(myTesting)


```

library(parallel)
library(doParallel)
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)
#develop Model

#stop parallel processing
stopCluster(cluster)


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
